### 2020-09-17 paper/coloringBergeGraphs.tex
Fixes

The polynomial technique of coloring perfect graphs known so far arose in the field of semidefinite programming. Semidefinite programs are linear programs over the cones of semidefinite matrices. The connection of coloring graphs and the cones of semidefinite matrices might be surprising, so let us take a brief digression into the field of information theory, where we will see the connection more clearly. Also, this was exactly the background which motivated Berge to introduce perfect graphs \cite{Chudnovsky2003}.

@@

Suppose we have a noisy communication channel in which certain signal values can be confused with others. For instance, suppose our channel has five discrete signal values, represented as 0, 1, 2, 3, 4. However, each value of $i$ when sent across the channel can be confused with value $(i \pm 1)$ mod $5$. This situation can be modeled by a graph $C_5$ (\cref{fig:c5}) in which vertices correspond to signal values and two vertices are connected if and only if values they represent can be confused.

@@

Let us be more precise. Given a graph $G$ modeling a communication channel and a number $k \geq 1$ we say that two messages $v_1v_2\ldots v_k$, $w_1w_2\ldots w_k \in V(G)^k$ of length $k$ are non-confoundable if and only if there is $1 \leq i \leq k$ such that $v_i$, $w_i$ are non-confoundable. We are interested in the maximum rate at which we can reliably transmit information (the \emph{Shannon capacity} of the channel defined by $G$).

For $k = 1$, maximum number of messages we can send without confusion in a single step is equal to $\alpha(G)$. To describe longer messages we use \emph{strong product} $G \cdot H$ of two graphs $G = (V, E)$, $H = (W, F)$ as the graph with $V(G \cdot H) = V \times W$, with $(i, u)(j, v) \in E(G \cdot H)$ if and only if $ij \in E$ and $uv \in F$, or $ij \in E$ and $u = v$, or $i = j$ and $uv \in F$. Given channel modeled by $G$ it is easy to see that the maximum number of distinguishable words of length 2 is equal to $\alpha(G \cdot G)$, and in general the number of distinguishable words of length $k$ is equal to $\alpha(G^k)$ -- which gives us $\sqrt[k]{\alpha(G^k)}$ as the number of distinguishable signals per single transmission. So, we can define the Shannon capacity of the channel defined by $G$ as $\Theta(G) = \sup\limits_k \sqrt[k]{\alpha(G^k)}$.

@@

If $S$ is any stable set in $C_5^k$, then $\{v_i: i \in S\}$ is a set of mutually orthogonal unit vectors, so we could extend $S$ to $S'$, so that $\{v_i: i \in S'\}$ is a basis of $\mathbb{R}^{3k}$. This gives us
$$\sum\limits_{i\in S}(e_1^\intercal v_i)^2 \leq \sum\limits_{i \in S'}(e_1^\intercal v_i)^2 \leq ? = 1$$
\todo{todo}
Each term on the left hand side is $5^{-k/4}$, so the left hand side is equal to $|S|5^{-k/2}$, and so $|S| \leq 5^{k/2}$. Since $|S|$ was an arbitrary stable set, we get $\alpha(C_5^k) \leq 5 ^{k/2}$ and $\Theta(C_5) = \sqrt{5}$.

@@

\begin{defn}[positive semidefinite matrix]
  Let $A$ be an $n \times n$ symmetric matrix. $A$ is \emph{positive semidefinite} if all of its eigenvalues are nonnegative. We denote it by $A \succeq 0$.
\end{defn}

@@

The fact that the sum of two positive semidefinite matrices is again positive semidefinite, with the fact that every positive scalar multiple of a positive semidefinite matrix is positive semidefinite, translates into the geometric statement that the set of all positive semidefinite matrices forms a convex closed cone $\mathcal{P}_n$ in $\mathbb{R}^{n \times n}$ with vertex 0. This cone $\mathcal{P}_n$ is important but its structure is not trivial.

@@

Any stable set $S$ of $G$ provides a feasible solution of the sual program, by choosing $Z_{ij} = \frac{1}{S}$, if $i, j \in S$ and 0 otherwise. Similarly, any $k$-coloring of $\overline{G}$ provides a feasible solution of the former semidefinite program, by choosing $Y_{ij} = -1$ if $i$ and $j$ have different colors, $Y_{ii} = k-1$, and $Y_{ij} = 0$ otherwise.

@@

\section{Coloring perfect graph using ellipsoid method}
\label{sec:coloringEllipsoid}
The following part is based on \citetitle{Laurent2005} by \citeauthor{Laurent2005} \cite{Laurent2005}.

@@

Given graph $G$, recall that stability number of $G$ is equal clique number of the complement of $G$. This gives us a way to compute $\alpha(G)$ for any perfect graph $G$.

@@

There are at most $|V|$ maximum cardinality cliques in $G$. Adding a single clique to the list of maximum cardinality cliques requires constructing auxiliary graph for weighted maximum stable set, which is of size $O(|V|^2)$ and running \Cref{alg:maxStableSet} on it. Therefore total running time is polynomial.

@@

Note that coloring $C'$ uses $\chi(G \setminus S) = \omega(G \setminus S) = \omega(G) -1$. We will call recursion at most $O(|V|)$ times, each step of recursion requires polynomial time. Therefore the total running time of \Cref{alg:minColoring} is polynomial.

@@

\begin{defn}[prism]
  A \emph{prism} is a graph consisting of two disjoint triangles and two disjoint paths between them. Notice that for a graph to contain no odd hole, all three paths in a prism must have the same parity. A prism with all three paths odd is called an \emph{odd prism}.
  \label{def:prism}
\end{defn}

In 2005 Maffray and Trotignon showed a coloring algorithm that colors graphs containing no odd hole, no antihole and no prism (sometimes called Artemis graphs) in $O(|V|^4|E|)$ time \cite{Maffray2006}. They later improved the time complexity to $O(|V|^2|E|)$ \cite{coloringArtemis}.

In 2015 Maffray showed an algorithm for coloring Berge graphs with no squares (a square is a $C_4$) and no odd prism \cite{Maff2015}.

@@

A most recent advancement (2018) is an algorithm by Chudnovsky et al. that colors any square-free Berge graphs in time of $O(|V|^9)$ \cite{coloringSquareFree}. Before proving strong perfect graph conjecture, a similar conjecture for square-free Berge graphs has been proven by Conforti et al. \cite{Conforti2004}. During one of her lectures, Maria Chudnovsky expressed hope that discovery of full algorithm for coloring Berge graphs might follow a similar pattern.

The algorithm for coloring square-free Berge graphs is much more complex than CCLSV algorithm for Berge graphs recognition and would be much harder to implement. The main idea is that, if square-free Berge graph $G$ is not a member of class that is simpler to analyze and color (Artemis graphs), then it must have an induced prism. We find it and grow it until it is maximal and then we can create a partition of $G$, that we color recursively. The paper in which the algorithm is presented \cite{coloringSquareFree} is of much more mathematical ''flavor'' than \cite{MC05} -- there is a proof that a $O(|V|^9)$ algorithm exists, but actually deducing it from the paper is far from straightforward. We provide its pseudocode in the Appendix A.

>>>>

The only polynomial technique of coloring perfect graphs known so far arose in the field of semidefinite programming. Semidefinite programs are linear programs over the cones of semidefinite matrices. The connection of coloring graphs and the cones of semidefinite matrices might be surprising, so let us take a brief digression into the field of information theory, where we will see the connection more clearly. Also, this was exactly the background which motivated Berge to introduce perfect graphs \cite{Chudnovsky2003}.

@@

Suppose we have a noisy communication channel in which certain signal values can be confused with others. For instance, suppose our channel has five discrete signal values, represented as 0, 1, 2, 3, 4. However, each value of $i$ when sent across the channel can be confused with value $(i \pm 1)$ mod $5$. This situation can be modeled by a graph $C_5$ (\cref{fig:c5}) in which vertices correspond to signal values and two vertices are connected if and only if the values they represent can be confused.

@@

Let us be more precise. Given a graph $G$ modeling a communication channel and a number $k \geq 1$ we say that two messages $v_1v_2\ldots v_k$, $w_1w_2\ldots w_k \in V(G)^k$ of length $k$ are non-confoundable if and only if there is $1 \leq i \leq k$ such that $v_i$, $w_i$ are non-confoundable. We are interested in the maximum rate at which we can reliably transmit information, that is, the \emph{Shannon capacity} of the channel defined by $G$.

For $k = 1$, maximum number of messages we can send without confusion in a single step is equal to $\alpha(G)$. To describe longer messages we use \emph{strong product} $G \cdot H$ of two graphs $G = (V, E)$, $H = (W, F)$ as the graph with $V(G \cdot H) = V \times W$, with $(i, u)(j, v) \in E(G \cdot H)$ if and only if $ij \in E$ and $uv \in F$, or $ij \in E$ and $u = v$, or $i = j$ and $uv \in F$. Given channel modeled by $G$ it is easy to see that the maximum number of distinguishable words of length 2 is equal to $\alpha(G \cdot G)$, and in general the number of distinguishable words of length $k$ is equal to $\alpha(G^k)$ which gives us $\sqrt[k]{\alpha(G^k)}$ as the number of distinguishable signals per single transmission. So, we can define the Shannon capacity of the channel defined by $G$ as $\Theta(G) = \sup\limits_k \sqrt[k]{\alpha(G^k)}$.

@@

If $S$ is any stable set in $C_5^k$, then $\{v_i: i \in S\}$ is a set of mutually orthogonal unit vectors, so we could extend $S$ to $S'$, so that $\{v_i: i \in S'\}$ is a basis of $\mathbb{R}^{3k}$. This gives us
$$\sum\limits_{i\in S}(e_1^\intercal v_i)^2 \leq \sum\limits_{i \in S'}(e_1^\intercal v_i)^2 \leq |e_1|^2 = 1$$
\todo{todo}
Each term on the left hand side is equal to $5^{-k/4}$, so the left hand side is equal to $|S|5^{-k/2}$, and so $|S| \leq 5^{k/2}$. Since $|S|$ was an arbitrary stable set, we get $\alpha(C_5^k) \leq 5 ^{k/2}$ and $\Theta(C_5) = \sqrt{5}$.

@@

\begin{defn}[positive semidefinite matrix]
  Let $A$ be an $n \times n$ symmetric matrix. $A$ is \emph{positive semidefinite} if and only if all of its eigenvalues are nonnegative. We denote it by $A \succeq 0$.
\end{defn}

@@

The fact that the sum of two positive semidefinite matrices is again positive semidefinite, with the fact that every positive scalar multiple of a positive semidefinite matrix is positive semidefinite, translates into the geometric statement that the set of all positive semidefinite matrices forms a convex closed cone $\mathcal{P}_n$ in $\mathbb{R}^{n \times n}$ with its apex at 0. This cone $\mathcal{P}_n$ is important but its structure is not trivial.

@@

Any stable set $S$ of $G$ provides a feasible solution of the dual program, by choosing $Z_{ij} = \frac{1}{S}$ if $i, j \in S$ and 0 otherwise. Similarly, any $k$-coloring of $\overline{G}$ provides a feasible solution of the former semidefinite program, by choosing $Y_{ij} = -1$ if $i$ and $j$ have different colors, $Y_{ii} = k-1$, and $Y_{ij} = 0$ otherwise.

@@

\section{Coloring perfect graphs using ellipsoid method}
\label{sec:coloringEllipsoid}
The following part is based on \citetitle{Laurent2005} by \citeauthor{Laurent2005} \cite{Laurent2005}.

@@

Given graph $G$, recall that the stability number of $G$ is equal the clique number of the complement of $G$. This gives us a way to compute $\alpha(G)$ for any perfect graph $G$.

@@

There are at most $|V|$ maximum cardinality cliques in $G$. Adding a single clique to the list of maximum cardinality cliques requires constructing auxiliary graph for weighted maximum stable set, which is of size $O(|V|^2)$ and running \Cref{alg:maxWeightedStableSet} on it. Therefore total running time is polynomial.

@@

Note that coloring $C'$ uses $\chi(G \setminus S) = \omega(G \setminus S) = \omega(G) -1$ colors, so we will call the recursion $O(|V|)$ times, each step of recursion requires polynomial time. Therefore the total running time of \Cref{alg:minColoring} is polynomial.

@@

\begin{defn}[prism]
  A \emph{prism} is a graph consisting of two disjoint triangles and two disjoint paths between them. Notice that for a graph to contain no odd holes, all three paths in a prism must have the same parity.
  \label{def:prism}
\end{defn}

A prism with all three paths odd is called an \emph{odd prism} and a prism with all three paths even is called an \emph{even prism}.

In 2005 Maffray and Trotignon showed a coloring algorithm that colors graphs containing no odd holes, no antiholes and no prisms (sometimes called Artemis graphs) in $O(|V|^4|E|)$ time \cite{Maffray2006}. They later improved the time complexity to $O(|V|^2|E|)$ \cite{coloringArtemis}. In 2015 Maffray showed an algorithm for coloring Berge graphs with no squares\footnote{sometimes called square-free Berge graphs} (a square is a $C_4$) and no odd prisms \cite{Maff2015}.

@@

A most recent advancement (2018) is an algorithm by Chudnovsky et al. that colors any square-free Berge graphs in time of $O(|V|^9)$ \cite{coloringSquareFree}. Before proving strong perfect graph conjecture a similar conjecture for square-free Berge graphs has been proven by Conforti et al. \cite{Conforti2004}. During one of her lectures, Maria Chudnovsky expressed hope that discovery of full algorithm for coloring Berge graphs might follow a similar pattern.

The algorithm for coloring square-free Berge graphs is much more complex than CCLSV algorithm for Berge graphs recognition and would be much harder to implement. The main idea is that, if square-free Berge graph $G$ is not a member of class that is simple to recognize and color (Artemis graphs), then it must have an induced prism. We find it and grow it until it is maximal and then we can create a partition of $G$, that we color recursively. The paper in which the algorithm is presented \cite{coloringSquareFree} is of much more mathematical ''flavor'' than \cite{MC05} -- there is a proof that a $O(|V|^9)$ algorithm exists, but actually deducing it from the paper is far from straightforward. We provide its pseudocode in the Appendix A.

### 2020-09-17 paper/definitions.tex
Fixes

An edge $\{x, y\}$ is said to \emph{join}, or \emph{be between} vertices $x$ and $y$ and is denoted by $xy$. Thus $xy$ and $yx$ denote the same edge (all our graphs are \emph{undirected}). If $xy \in E(G)$ then $x$ and $y$ are \emph{adjacent}, \emph{connected} or \emph{neighboring}. By $N(x)$ we will denote the \emph{neighborhood} of $x$, that is, all vertices $y$ such that $xy$ is an edge. Similarly, for $X \subseteq V(G)$, by $N(X)$ we will denote the \emph{neighborhood} of $X$, meaning all vertices of $v \in V(G) \setminus X$, so that there is a $x \in X$, that $xv$ is an edge in $G$. If $xy \notin E(G)$ then $xy$ is a \emph{nonedge} and $x$ and $y$ are \emph{anticonnected}.

@@

\begin{defn}[subgraph]
  $G' = (V', E')$ is a \emph{subgraph} of $G = (V, E)$ if and only if $V' \subseteq V$ and $E' \subseteq E$.
\end{defn}

\begin{defn}[induced subgraph]
  If $G' = (V', E')$ is a subgraph of $G$ and it contains \emph{all edges} of $G$ that join two vertices in $V'$, then $G'$ is said to be \emph{induced subgraph} of $G$ and is denoted $G[V']$.
\end{defn}

@@

Graph $G_0$ is a path of length 3, with the inside $G_0^* = \{v_2, v_3\}$. If we  added any edge to $G_0$ it would stop being a path.

@@

This cycle $C$ is usually denoted by $x_1x_2\ldots x_lx_1$ or $x_1 - x_2 - \ldots - x_l - x_1$. $l = |E(C)|$ is the \emph{length} of the cycle $C$. Sometimes we will denote the cycle of length $l$ as $C_l$.

@@

\begin{defn}[antihole]
  An \emph{antihole} is a subgraph of $G$ whose complement is a hole in $G$.
\end{defn}

>>>>

An edge $\{x, y\}$ is said to \emph{join}, or \emph{be between} vertices $x$ and $y$ and is denoted by $xy$. Thus $xy$ and $yx$ denote the same edge (all our graphs are \emph{undirected}). If $xy \in E(G)$ then $x$ and $y$ are \emph{adjacent}, \emph{connected} or \emph{neighboring}. By $N(x)$ we will denote the \emph{neighborhood} of $x$, that is, all vertices $y$ such that $xy$ is an edge. Similarly, for $X \subseteq V(G)$, by $N(X)$ we will denote the \emph{neighborhood} of $X$, meaning all vertices of $v \in V(G) \setminus X$, so that there is a $x \in X$, that $xv$ is an edge in $G$. If $xy \notin E(G)$ then $xy$ is a \emph{nonedge} and $x$ and $y$ are \emph{nonneighbors} or \emph{anticonnected}.

@@

\begin{defn}[subgraph]
  A graph $G' = (V', E')$ is a \emph{subgraph} of $G = (V, E)$ if and only if $V' \subseteq V$ and $E' \subseteq E$.
\end{defn}

\begin{defn}[induced subgraph]
  If $G' = (V', E')$ is a subgraph of $G$ and it contains \emph{all edges} of $G$ that join two vertices in $V'$, then $G'$ is said to be an \emph{induced subgraph} of $G$ and is denoted $G[V']$.
\end{defn}

@@

Our graph $G_0$ is a path of length 3, with the inside $G_0^* = \{v_2, v_3\}$. If we  added any edge to $G_0$ it would stop being a path.

@@

This cycle $C$ is usually denoted by $x_1x_2\ldots x_lx_1$ or $x_1$-$x_2$-$\ldots$-$x_l$-$x_1$. $l = |E(C)|$ is the \emph{length} of the cycle $C$. Sometimes we will denote the cycle of length $l$ as $C_l$.

@@

\begin{defn}[antipath]
  An \emph{antipath} is a graph $G$ such that $\overline{G}$ is a path.
\end{defn}

\begin{defn}[antihole]
  An \emph{antihole} is a graph $G$ such that $\overline{G}$ is a hole.
\end{defn}

### 2020-09-17 paper/experiments.tex
Fixes

Let us describe experiments and their results. All of our algorithms search for an odd hole or an odd antihole and stop when find one, or a evidence of one. Therefore, their running times are greatest with perfect graphs on the input and we decided to use\todo{run some smart nonperfect tests} perfect graphs for our performance benchmarks. For every test, the vertex number were shuffled. For every randomly generated graph class and for each size, we ran the tests on 10 graphs generated with different seeds and took their average running time.

Below, we first present our data sets then overall time results on them and lastly a plot breaking down what parts of CCLSV and GPU CCLSV constitute their overall running time.

@@

\todo{generate perfect Erdos--Renyi graphs}

@@

% An \emph{Erd\H{o}s--Rényi} model is a method of generating random graphs. Every possible edge is included in a graph with probability $1/2$. We generated them as long as we found enough perfect graphs for each $|V|$ we wanted to test. \todo{implement generating perfect graphs by removing long cycles} This method of generating perfect graphs is very inefficient and we couldn't generate any graphs with $|V| \geq 20$. Second, those graphs favour the na\"ive algortihm, because there is very low probability of long paths to appear.

We generate random bipartite graphs by creating two equal-sized sets of vertices, and then added every possible edge between them with a probability of $p = 0.5$. In this way we can also generate line graphs of random bipartite graphs. This gives us graphs of varying $|V|$, so we repeated the process until we had sufficieng number of graphs for each $|V|$.

There are also a few interesting classes of perfect graphs that appear on a grid, or a checker board. In lattice graph, each vertex is connected to vertices that are above, below, to the left and to the right of it, if such vertices exist. See \Cref{fig:lattice} for an example of a $4 \times 4$ lattice graph. If we imagine our grid to be a chessboard we can define two vertices a rook graph to share an edge, when the rook from chess can move from one vertex to the other in one move. See \Cref{fig:rook} for an example of $4 \times 4$ rook graph. In similar manner we can define a knight graph (\Cref{fig:knight}).

@@

First, let us take a look at the performance of CCLSV algorithm. If we limit runtime to about two minutes, we can run it on graphs of sizes up to 40-70, depending on the type of the graph we test.  Worth noting is the fact that there isn't that much of a variance of the running time depending on the type of graph we test on. This is a big advantage of CCLSV over the na\"ive algorithm, which is highly dependent on the type of input. On some graphs, it works almost instantenously, what is to be expected, given its method of enumerating paths. But on some graphs, e.g. lattice graphs (\Cref{plot:gridLines}) its overall running time grows much faster than CCLSV's.

Let us turn our attention to the GPU CCLSV's results. \Cref{plot:perfDetailed} shows breakdown of each implementation running times on a single graph of each class. We can clearly see that in most cases testing possible near-cleaners takes most of the running time, so the speedup by the GPU is significant.

Examples of full binary graphs, and lattice graphs show that getting possible near cleaners still can be a big bottleneck. Parallelization of this method could be considered in the future. A possible idea would be to use recently developed dynamic dictionary for the GPU \cite{Ashkiani2018}. When running GPU CCLSV on split graphs, quite a lot of time is spent checking simple structures. The biggest challange we can see is in an attempt to parallelise this is spreading work evenly across CUDA threads.

% When looking at the results (\Cref{plot:perfLines}), while na\"ive algorithm's time is almost zero, we see the CCLSV running time climbing with the growth of $|V|$, with only moderate improvement by GPU CCLSV. Slight GPU improvement is explained by anylysis of the components of the overall time (\Cref{plot:perfDet}). Testing all near cleaners takes up around half of total time, so we cannot speed up the overall time by much. These are also relatively small tests, so the latency from copying data to the GPU is significant.

% Next, we generated a random bipartite graphs (see paragraph above) and then calculated their line graphs. This gives us graphs of varying $|V|$, so we repeated the process until we had sufficieng number of graphs for each $|V|$. On \cref{plot:perf2Lines} we can see na\"ive algorithm being much faster on the majority of the tests, although its faster growth allows GPU CCLSV to be faster on the biggest tests.

@@

Calculating $\vartheta(G)$ is the most complicated part of the coloring algorithm. With that done by an external library, the rest of the program is a straightforward implementation of the algorithms in \Cref{sec:coloringEllipsoid}. In most of our tests, the majority of running time was consumed on calculating $\vartheta$ of various graphs. As it was done by an external library, there isn't much optimization potential for us. One thing of note is, that by specifying the precision of $\vartheta$ we want to calculate to be $1/3$ the algorithm sped up  by 20-40\%. Also, we cache results of calculatin $\vartheta$, meaning that throught whole execution of our coloring algorithm, we only calculate $\vartheta$ for each graph once. Our main goal of the implementation was to check if this method is still impractical, even on modern equipment. Let us proceed straight to experiments and results.

@@

We ran our tests on the data sets generated in the same way as for perfect graph recognition. We only adjusted grah sizes to get the running times of up to a couple of minutes.

@@

First, we present overall running times of the coloring algorithm (\Cref{fig:CSDPLines}) and then breakdown the running time in two parts: calculating $\vartheta$ by the CSDP library and graphs manipulations which constitute the rest of the coloring algorithm (\Cref{fig:CSDPDet}). We note that $\vartheta$ calculation is highly dependent on the type of the graph. For our biggest tested cases of $|V| = 48$, all required calculations of $\vartheta$ take almost 7 minutes in bipartite graphs, but only 12 seconds in line graphs of bipartite graphs. Because CSDP library is highly optimized, it is expected that it would run much faster on some graphs. However, it is not clear what could we do to speedup the running times on tests that CSDP is slow on, such as bipartite graphs or lattice graphs. A potential improvement would be to again use CUDA, this time for $\vartheta$ calculations, possibly utilizing the cuSOLVER library \cite{cusolver}.

@@

\section{Conclusions}
We implemented and benchmarked two different approaches for recognizing perfect graphs. Both have their positives and drawbacks -- CCLSV's running times are more predictable, but on many graphs it is noticeably slower than the na\"ive algorithm. On the other hand, running na\"ive algorithm on a graph we don't know any properties of can lead to unbearably slow performance.

The GPU CCLSV implementation shows that using parallel computation can be very profitable for this task. A further research can be done to see if further progress could be made, however there is no obvious improvement we can see. Generating all possible near-cleaners makes a heavy use of sets (our tests indicated that if we simply list all combinations of sets $X_i$ and $N_j$, the resulting list $\mathcal{R}$ would be much bigger than necessary). Possibly a newly developed dynamic dictionary for the GPU \cite{Ashkiani2018} would be of use here. Testing for simple forbidden structures should also be parallelizable, although our preliminary tests indicated that spreading all the work evenly between CUDA threads is not trivial.

A GPU implementation of na\"ive algorithm could also be considered for future work. We have created one, but after all path-enumaration optimizations of the CPU implementation we didn't manage to achieve any noticeable speedup. The reason again was, that we could't spread the work evenly between CUDA threads. Because some prefixes of paths require much more processing than others, this is not a trivial task and a careful study would be needed.

We also implemented algorithm for coloring perfect graphs. It is not entirely unusable, although far from fast. Maybe a better way of calculating $\vartheta$ could be found, or the GPU could again be usable for this task. We point to Nvidia's cuSOLVER  cite{cusolver} as a possible starting point.

>>>>

Let us describe experiments and their results. All of our algorithms search for an odd hole or an odd antihole and stop when find one, or a evidence of one. Therefore, their running times are greatest with perfect graphs on the input and we decided to use perfect graphs for our performance benchmarks. For every test, the vertex numbers were shuffled. For every randomly generated graph class and for each size, we ran the tests on 10 graphs generated with different seeds and took their average running time.
% \todo{run some smart nonperfect tests}

Below, we first present our data sets then overall time results on them and lastly a plot breaking down what parts of CCLSV and GPU CCLSV implementations constitute their overall running times.

@@

% \todo{generate perfect Erdos--Renyi graphs}

@@

% An \emph{Erd\H{o}s--Rényi} model is a method of generating random graphs. Every possible edge is included in a graph with probability $1/2$. We generated them as long as we found enough perfect graphs for each $|V|$ we wanted to test. \todo{implement generating perfect graphs by removing long cycles} This method of generating perfect graphs is very inefficient and we couldn't generate any graphs with $|V| \geq 20$. Second, those graphs favor the na\"ive algorithm, because there is very low probability of long paths to appear.

We generate random bipartite graphs by creating two equal-sized sets of vertices, and then adding every possible edge between them with a probability of $p = 0.5$. In this way we can also generate line graphs of random bipartite graphs. This gives us graphs of varying $|V|$, so we repeated the process until we had sufficient number of graphs for each $|V|$.

There are also a few interesting classes of perfect graphs that appear on a grid, or a chessboard. In lattice graph, each vertex is connected to vertices that are above, below, to the left and to the right of it, if such vertices exist. See \Cref{fig:lattice} for an example of a $4 \times 4$ lattice graph. If we imagine our grid to be a chessboard we can define two vertices a rook graph to share an edge, when the rook from chess can move from one vertex to the other in one move. See \Cref{fig:rook} for an example of $4 \times 4$ rook graph. In similar manner we can define a knight graph (\Cref{fig:knight}).

@@

First, let us take a look at the performance of the CCLSV algorithm. If we limit the runtime to about two minutes, we can run it on graphs of sizes up to 40-70, depending on the type of the graph we test. Worth noting is the fact that the CCLSV algorithm usually grows slower than the na\"ive algorithm. Oftentimes this is mitigated by the fact that the na\"ive algorithm works very well on smaller graphs, but it is not always the case. For example, for larger instances of lattice graphs and hypercube graphs the CCLSV works better than na\"ive algorithm. If we would want to test graphs larger than presented here, the GPU CCLSV seems to be the best solution. For every type of graph we tested, it finished within two minutes for graphs of $|V|$ up to about 50. The na\"ive algorithm is very fast for some graphs, like split graphs or random bipartite graphs, but GPU CCLSV still seems to be the best solution for larger graphs of unknown structure.
  % there isn't that much of a variance of the running time depending on the type of graph we test on. This is a big advantage of CCLSV over the na\"ive algorithm, which is highly dependent on the type of input. On some graphs, it works almost instantaneously, what is to be expected, given its method of enumerating paths. But on some graphs, e.g. lattice graphs (\Cref{plot:gridLines}) its overall running time grows much faster than CCLSV's.

Let us try to understand where the speedup of GPU CCLSV comes from. \Cref{plot:perfDetailed} shows breakdown of each implementation running times on a single graph of each class. We can clearly see that in most cases testing possible near-cleaners takes most of the running time, so the speedup achieved by utilizing the GPU is significant.

Examples of full binary graphs, and lattice graphs show that getting possible near cleaners still can be a big bottleneck. Parallelization of this method could be considered in the future. A possible idea would be to use recently developed dynamic dictionary for the GPU \cite{Ashkiani2018}. When running GPU CCLSV on split graphs, quite a lot of time is spent checking simple structures. The biggest challenge we can see is in an attempt to parallelize this is spreading work evenly across CUDA threads.

% When looking at the results (\Cref{plot:perfLines}), while na\"ive algorithm's time is almost zero, we see the CCLSV running time climbing with the growth of $|V|$, with only moderate improvement by GPU CCLSV. Slight GPU improvement is explained by analysis of the components of the overall time (\Cref{plot:perfDet}). Testing all near cleaners takes up around half of total time, so we cannot speed up the overall time by much. These are also relatively small tests, so the latency from copying data to the GPU is significant.

% Next, we generated a random bipartite graphs (see paragraph above) and then calculated their line graphs. This gives us graphs of varying $|V|$, so we repeated the process until we had sufficient number of graphs for each $|V|$. On \cref{plot:perf2Lines} we can see na\"ive algorithm being much faster on the majority of the tests, although its faster growth allows GPU CCLSV to be faster on the biggest tests.

@@

Calculating $\vartheta(G)$ is the most complicated part of the coloring algorithm. With that done by an external library, the rest of the program is a straightforward implementation of the algorithms in \Cref{sec:coloringEllipsoid}. In most of our tests, the majority of running time was consumed on calculating $\vartheta$ of various graphs. As it was done by an external library, there isn't much optimization potential for us. One thing of note is, that by specifying the precision of $\vartheta$ we want to calculate to be $1/3$ the algorithm sped up  by 20-40\%. Also, we cache results of calculating $\vartheta$, meaning that through whole execution of our coloring algorithm, we only calculate $\vartheta$ for each graph once. Our main goal of the implementation was to check if this method is still impractical, even on modern equipment. Let us proceed straight to experiments and results.

@@

We ran our tests on the data sets generated in the same way as for perfect graph recognition. We only adjusted graph sizes to get the running times of up to a couple of minutes.

@@

First, we present overall running times of the coloring algorithm (\Cref{fig:CSDPLines}) and then breakdown the running time in two parts: calculating $\vartheta$ by the CSDP library and graphs manipulations which constitute the rest of the coloring algorithm (\Cref{fig:CSDPDet}). We note that $\vartheta$ calculation is very dependent on the type of the graph. For our biggest tested cases of $|V| = 48$, all required calculations of $\vartheta$ take almost 7 minutes in bipartite graphs, but only 12 seconds in line graphs of bipartite graphs. Because CSDP library is highly optimized, it is expected that it would run much faster on some graphs. However, it is not clear what we could do to speedup the running times on tests that CSDP is slow on, such as bipartite graphs or lattice graphs. A potential improvement would be to again use CUDA, this time for $\vartheta$ calculations, possibly utilizing the cuSOLVER library \cite{cusolver}.

@@

\section{Conclusions}
We implemented and benchmarked two different approaches for recognizing perfect graphs. Both have their positives and drawbacks -- testing big graph of unknown structure is probably better done with the CCLSV algorithm, while the na\"ive algorithm is much faster on some cases. The GPU CCLSV implementation shows that using parallel computation can be very profitable for this task. An additional research could be done to see if further progress can be made, however there is no obvious improvement we can see. Generating all possible near-cleaners makes a heavy use of sets (our tests indicated that if we simply list all combinations of sets $X_i$ and $N_j$, the resulting list $\mathcal{R}$ would be much bigger than necessary). Possibly a newly developed dynamic dictionary for the GPU \cite{Ashkiani2018} would be of use here. Testing for simple forbidden structures should also be parallelizable, although our preliminary tests indicated that spreading all the work evenly between CUDA threads is not trivial.

A GPU implementation of the na\"ive algorithm could also be considered for future work. We have created one, but after all path-enumeration optimizations of the CPU implementation we didn't manage to achieve any noticeable speedup. The reason again was, that we couldn't spread the work evenly between CUDA threads. Because some prefixes of paths require much more processing than others, this is not a trivial task and a careful study would be needed.

We also implemented algorithm for coloring perfect graphs. It is not entirely unusable, although far from fast. Maybe a better way of calculating $\vartheta$ could be found, or the GPU could again be usable for this task. We point to Nvidia's cuSOLVER  \cite{cusolver} as a possible starting point.
